---
title: "Getting Familiar with ncov2019"
author: "Silas Morsink and Matthew Colon"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Familiar with ncov2019}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Importing Data

To easily import data on COVID-19, Zika, or SARS, use the corresponding `import` function:

``` {r}
covid_data <- importCovidData()    # default from_web = TRUE
zika_data <- importZikaData()      # default from_web = FALSE
sars_data <- importSARSData()      # default from_web = FALSE
```

Because Zika data and SARS data are no longer changing, it is not necessary to load data from the web. By setting `from_web = FALSE`, the function simply loads the data from the `data` directory. For COVID-19, to get the most up-to-date information, we need `from_web = TRUE` (the default). If you set `from_web = FALSE` for `importCovidData()`, the function will load the saved COVID-19 data from the `data` directory, which only goes through 2020-03-06. 

Behind the scenes, the `import` functions are using other functions for data acquisition and cleaning. These functions are also exported for the user to observe and potentially modify. `importCovidData()` calls `scrapeCovidData()` to pull data from the web, then `tidyCovidData()` to reformat it, then `accumulateCovidData()` to pull together data on cases, deaths, and recovered.

