---
title: "Report"
author: "Matthew Colón"
date: "3/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package Description

This package drastically improves the user's ability to import, filter, augment, and visualize disease data from the 2003 SARS virus, the 2015-16 Zika, and the 2020 Coronavirus.


## Use Cases for our Package

You can also embed plots, for example:

```{r pressure, echo=FALSE}
```

## Team Member Contributions

### Silas Morsink
Silas was responsible for the following:

* Data Import/Cleaning
    + Wrote functions to import, clean, and store Coronavirus data
    + Wrote functions to import, clean, and store Zika data
    + Wrote functions to import and store SARS data
* Data Filtering/Augmentation
    + Wrote the buildCoordinateMap function
    + Wrote the buildPopulationMap function
* Data Visualization
    + Wrote the plotTimeSeries function
* Package Documentation
    + Wrote the package README file
    + Wrote the package vignette

### Matthew Colón
Matthew was responsible for the following:

* Data Import/Cleaning
    + Cleaned the SARS data to be used in the package
* Data Filtering/Augmentation
    + Wrote the filterDiseaseData function
    + Wrote the importCoordinateData function
    + Wrote the dayOfDiseaseColumn function
* Data Visualization
    + Wrote the mapPlotStatic function
    + Wrote the mapPlotAnimate function
* Documentation
    + Wrote the package project Report

